A new type of generic, self-evolving and efficient automated deduction algorithm based on category theory

### 2023-04-25 toTranslate.tex
restart dev

\begin{document}
	\title{A new type of generalized, self-evolving and efficient automated deduction algorithm based on category theory}
	\author{%
		Shao Xinhui\affil{1}\corrauth
		Zijian Wang\affil{2}
	}
	\shortauthors{the Authors}
	\address{%
		\addr{\affilnum{1}}{Department of Mathematics, College of Sciences, Northerneastern University, Shenyang, Liaoning, P.R.China}
		\addr{\affilnum{2}}{Northeast Yucai School, Zhenxing St. 9-5, Heping District, Shenyang, Liaoning, P.R.China}
		\corraddr{shaoxinhui@mail.neu.edu.cn}
	}

	\begin{abstract}
	In this article a new type of generalized, self-evolving and efficient automated statement proof algorithm based on new data structures, i.e. brackets and map graphs
	and new algorithms is presented. The brackets structure provides an elegant low-knowledge representation of mathematical concepts. The map graphs offer an efficient machine-learning method which let the computer learn knowledge while proving. Additionally, the new finding is totally built on the category theory. Furthermore, a prototype of the program is given and examined for performance.
	\end{abstract}

@@

	\textit{ Structure of the paper.}\newline
	The paper mainly consists of seven sections. The Introduction section demonstrates the background of the research and fundamental opinions around it. The Preliminaries section, made up of two subsections, claims to minimized pre-knowledge required to understand the paper and the symbols we select to use. The third section defines a set of mathematical structures that will be used by the algorithm. The forth section explains the detailed workflow procedures of the algorithm. The next section claims the conclusions of the research and the to-do improvements in the future. Finally the Acknowledgments section expresses my gratitude to the help and the References section claims the referential resources used while conducting the project.

@@

	First, we need to observe what notation is needed by the construction. Apparently, we need to at least define these notations: \(\text{inf}\), \(\text{int}\), \(\text{num}\),\textit{  }\textit{ \(\sum\)}, \(=\), \({}^{\wedge}\), \(/\), \(-\), \(\text{abs}\), \(<\) and \(\text{for}\), where \(\text{abs}\) stands for the absolute value function. However, in this neat example, the concept of function is not needed to be defined. Additionally, as the constructed stuff only needs to demonstrate the equation itself instead of the principles behind the equality, there is no necessity for the definition of notation \(+\). Note that manifestly \(\text{int} \triangleleft | \text{num}\). The table below (\ref{table-1}) demonstrates the number of parameters taken by the notations.

	\begin{table}[h]
		\centering
		\caption{The table for $\mathit{p}(\mathcal{N})$}
		\begin{tabular}{ccccccccccc}
			\toprule
			$\mathcal{N}$ & int & num & inf & $\sum$ & $=$ & $/$ & - & abs & $<$ & for \\
			\midrule
			$\mathit{p}$ & $0$ & $0$ & $0$ & $3$ & $2$ & $2$ & $2$ & $1$ & $2$ & $2$ \\
			\bottomrule
		\end{tabular}
		\label{table-1}
	\end{table}

@@

	\begin{equation*}
		\begin{aligned}
			\pmb{1} &\in& \text{Ob}(\text{Sym}(\mathcal{C}))\land \mathit{v}(\pmb{1})=\text{number}\\
			\pmb{\infty } &\in& \text{Ob}(\text{Sym}(\mathcal{C}))\land \mathit{v}(\pmb{\infty })=\text{infinity}\\
			\pmb{x} &\in& \text{Ob}(\text{Sym}(\mathcal{C}))\land \mathit{v}(\pmb{x})=\text{number}\\
			\pmb{n} &\in& \text{Ob}(\text{Sym}(\mathcal{C}))\land \mathit{v}(\pmb{n})=\text{integer}\triangleleft |\text{number}\\
		\end{aligned}
	\end{equation*}
	\begin{equation*}
		\begin{aligned}
			\alpha &=& \{(0,\text{for}),(1,\beta ),(2,\gamma )\}\\
			\beta &=& \{(0,=),(1,\delta ),(2,\epsilon )\}\\
			\gamma &=& \{(0,<),(1,\zeta ),(2,\pmb{1})\}\\
			\delta &=& \{(0,\sum ),(1,\eta ),(2,\pmb{\infty }), (3,\theta )\}\\
			\epsilon &=& \{(0,/),(1,\pmb{1}),(2,\iota )\}\\
			\zeta &=& \{(0,\text{abs}),(1,\pmb{x})\}\\
			\eta &=& \{(0,=),(1,\pmb{n}),(2,\pmb{1})\}\\
			\theta &=& \{(0,{}^{\wedge}),(1,\pmb{x}),(2,\pmb{n})\}\\
			\iota &=& \{(0,-),(1,\pmb{1}),(2,\pmb{x})\}
		\end{aligned}
	\end{equation*}

@@

	\begin{equation}
		\begin{split}
			\alpha = \{(0,\text{for}),(1,\beta ),(2,\{(0,<),(1,\{(0,\text{abs}),\\(1,\pmb{x})\}),(2,\pmb{1})\})\}\hfill\\
			\beta = \{(0,=),(1,\delta ),(2,\{(0,/),(1,\pmb{1}),(2,\{(0,-),\\(1,\pmb{1}),(2,\pmb{x})\})\})\}\hfill\\
			\delta = \{(0,\sum ),(1,\{(0,=),(1,\pmb{n}),(2,\pmb{1})\}),(2,\pmb{\infty }), (3,\\\{(0,{}^{\wedge}),(1,\pmb{x}),(2,\pmb{n})\})\}\hfill\\
		\end{split}
	\end{equation}

@@

		\begin{equation}
			\begin{gathered}
				\mathit{e}\mathit{v}\mathit{a}\ell :\text{Ob}(\text{Mic}(\mathcal{C}))\to \mathcal{B}\\
				\mathit{m}\overset{\mathit{e}\mathit{v}\mathit{a}\ell }{\mapsto }\mathit{e}\mathit{v}\mathit{a}\ell (\mathit{m})
			\end{gathered}
		\end{equation}

@@

		Graphically, figure \ref{figure-1} illustrates the initial map.

		\begin{figure}[ht]
			\centering
			\begin{tikzpicture}[commutative diagrams/every diagram]
				\node (K0) at (90:2.3cm) {$\mathcal{K}_0$};
				\node (K1) at (90+72:2cm) {$\mathcal{K}_1$};
				\node (K2) at (90+2*72:2cm) {$\mathcal{K}_2$};
				\node (K3) at (90+3*72:2cm) {$\mathcal{K}_3$};
				\node (Kn) at (90+4*72:2cm) {$\mathcal{K}_N$};

				\path[commutative diagrams/.cd, every arrow, every label]
				(K0) edge node {$k_0$} (K1)
				(K1) edge node {$K_1$} (K2)
				(K2) edge node {$k_2$} (K3);

				\path[commutative diagrams/.cd, every arrow, every label, dashed]
				(K3) edge node {$K_3$} (Kn)
				(Kn) edge node {$k_N$} (K0);
			\end{tikzpicture}
			\caption{The initial map.}
			\label{figure-1}
		\end{figure}

@@

	\begin{algorithm}[h!]
		\caption{Pseudo code for recursive functor $\mathcal{F}_t$}
		\begin{algorithmic}[1]
			\Require Pool $\mathcal{M}_{t}$, Knowledge base $\mathcal{Q}$, random $\mathit{r}$, Just used reflection functor index $i$
			\Ensure Next pool $\mathcal{M}_{t+1}$

			\Function{$\mathcal{F}_t$}{$\mathcal{M}_t$, $\mathcal{Q}$, $\mathit{r}$, $i$}
			\Comment{This is the sub-process}
			\State $\mathcal{K} \gets $ \Call{SelectReflectionFunctor}{$\mathcal{Q}$, $\mathit{r}$, $i$}
			%% \State \Return \Call{Reflect}{$\mathcal{K}$, $\mathcal{M}_{t}$}
			\State \Return $\mathcal{K} \mathcal{M}_t$
			\Comment{Apply functor $\mathcal{K}$ to pool}
			\EndFunction

			\Function{SelectReflectionFunctor}{$\mathcal{Q}$, $\mathit{r}$, $i$}
			\Comment{Just used: $\mathcal{K}_i \in \text{Ob(}\mathcal{Q}\text{)}$}
			\State $M \gets Mor(\mathcal{Q})$
			\ForAll{$m \in M$}
			\Comment{Get morphisms starting at $\mathcal{K}_i$}
			\If{$s(m) != \mathcal{K}_i$}
			\State $M \gets M \backslash \{m\}$
			\EndIf
			\EndFor
			\If{$|M|==1$}
			\Comment{The easiest condition}
			\State \Return $t(m \in M)$
			\Else
			\State $n \gets \mathit{r} \% |M|$
			\State \Return $t(n)$
			\EndIf
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

@@

	A situation after some reinforcements may be the one demonstrated in figure \ref{figure-2}.

	\begin{figure}[h!t]
		\centering
		\begin{tikzpicture}[commutative diagrams/every diagram]
			\node (K0) at (90:2.3cm) {$\mathcal{K}_0$};
			\node (K1) at (90+72:2cm) {$\mathcal{K}_1$};
			\node (K2) at (90+2*72:2cm) {$\mathcal{K}_2$};
			\node (K3) at (90+3*72:2cm) {$\mathcal{K}_3$};
			\node (Kn) at (90+4*72:2cm) {$\mathcal{K}_N$};

			\path[commutative diagrams/.cd, every arrow, every label]
			(K0) edge node {$k_0$} (K1)
			(K1) edge node {$K_1$} (K2)
			(K2) edge node {$k_2$} (K3)
			(K1) edge node {} (K3)
			(K1) edge node {} (Kn)
			(K0) edge node {} (K3);

			\path[commutative diagrams/.cd, every arrow, every label, dashed]
			(K3) edge node {$K_3$} (Kn)
			(Kn) edge node {} (K1)
			(Kn) edge node {$k_N$} (K0);
		\end{tikzpicture}
		\caption{An example map after several reinforcements.}
		\label{figure-2}
	\end{figure}

@@

	To summarize, we can illustrate the entire process of the workflow of the algorithm with the following diagram on the next page (figure \ref{figure-3}).

	\begin{figure}[ht]
		\centering
		\begin{tikzpicture}[>=latex', font={\sf \small}]
			\def\smbwd{1cm}

			\node (start) at (0,0) [draw, terminal,
			minimum width = \smbwd,
			minimum height = 0.5cm] {Start};

			\node (parse) at (0,-1.5) [draw, process,
			minimum width = \smbwd,
			minimum height = 0.5cm] {Get task from user};

			\node (l-pool) at (-2,-3) [draw, storage,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$\mathcal{M}_t$};

			\node (r-pool) at (2,-3) [draw, storage,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$\mathcal{R}$};

			\node (rec-functor) at (-2,-4.5) [draw, predproc,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$\mathcal{F}_t$};

			\node (random) at (-4,-4.5) [draw, process,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$\mathit{r}$};

			\node (k-base) at (2,-4.5) [draw, storage,
			minimum width = \smbwd,
			minimum height = 0.5cm] {KBase $\mathcal{Q}$};

			\node (success-val) at (0,-6) [draw, process,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$s = {\exists n\in [0,\omega ), \mathcal{R}\subseteq \text{Ob}\left(\mathcal{M}_n\right)}$};

			\node (success-check) at (0,-7.5) [draw, decision,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$s$};

			\node (reinforcement) at (2,-9) [draw, predproc,
			minimum width = \smbwd,
			minimum height = 0.5cm] {Reinforcement};

			\node (write-rh) at (-2,-9) [draw, process,
			minimum width = \smbwd,
			minimum height = 0.5cm] {Generate steps};

			\node (steps) at (-4,-9) [draw, storage,
			minimum width = \smbwd,
			minimum height = 0.5cm] {Steps};

			\node (forgetful-para) at (2,-7.5) [draw, process,
			minimum width = \smbwd,
			minimum height = 0.5cm] {$E(\mathit{t})$};

			\node (end) at (0,-10.5) [draw, terminal,
			minimum width = \smbwd,
			minimum height = 0.5cm] {End};

			\coordinate (point1) at (-4.8,-7.5);
			\coordinate (point2) at (3.2,-4.5);
			\coordinate (point3) at (3.2,-8);
			%% \coordinate (point2) at (-7,-3);

			\draw[->] (start) -- (parse);
			\draw[->] (parse) |- (l-pool);
			\draw[->] (parse) |- (r-pool);
			\draw[->] (random) -- (rec-functor);
			\draw[->] (l-pool) -- (rec-functor);
			\draw[->] (k-base) -- (rec-functor);
			\draw[->] (l-pool) -| (success-val);
			\draw[->] (r-pool) -| (success-val);
			\draw[->] (success-val) -- (success-check);
			%% \draw[->] (success-check) -| node[above]{No} (l-pool);
			\draw[->] (success-check) -- node[above]{No} (point1);
			\draw[->] (point1) |- (l-pool);
			\draw[->] (success-check) |- node[below]{Yes} (write-rh);
			\draw[->] (success-check) |- node[below]{Yes} (reinforcement);
			\draw[->] (write-rh) -- (steps);
			\draw[->] (write-rh) |- (end);
			\draw[->] (forgetful-para) -- (reinforcement);
			%% \draw[->] (reinforcement) -- (k-base);
			\draw[->] (point2) -- (k-base);
			\draw[->] (point3) -- (point2);
			\draw[->] (reinforcement) |- (point3);
			\draw[->] (reinforcement) |- (end);
		\end{tikzpicture}
		\caption{The flowchart of the entire algorithm.}
		\label{figure-3}
	\end{figure}

@@

@@

	In the sections before we have presented an elegant data structure for representing mathematical concepts and a neat process which utilizes category theory concepts and knowledge to conduct proof, which also have self-optimiza-tion features.

@@

	\section{Acknowledgments}
	I express my sincerest gratitude to my tutor Prof. Shao Xinhui, the Yingcai Project of P.R.China, which enables senior high school students to conduct science research, my senior high academy Northeast Yucai School (NEYC) of P.R.China and the corresponding university Northern Eastern University (NEU) of P.R.China. I also thank my parents for encouraging me to conduct this research project.

@@

%	\bibliography{paper}
	\bibliographystyle{unsrtnat}
	\bibliography{paper}
\end{document}

>>>>

\begin{document}
	\title{A new type of generic, self-evolving and efficient automated deduction algorithm based on category theory}

    The authors' names are deleted.
    \author{%
    	Z\affil{1}
    	S\affil{*}
    }
    \shortauthors{the Authors}
    \address{%
    	\addr{\affilnum{1}}{N}
    	\corraddr{D}
    }

	\begin{abstract}
		Abstract deleted for privacy.
	\end{abstract}

@@

	\textit{ Structure of the paper.}\newline
	The paper mainly consists of seven sections. The Introduction section demonstrates the background of the research and fundamental opinions around it. The Preliminaries section, made up of two subsections, claims to minimized pre-knowledge required to understand the paper and the symbols we select to use. The third section defines a set of mathematical structures that will be used by the algorithm. The forth section explains the detailed workflow procedures of the algorithm. The next section claims the conclusions of the research and the to-do improvements in the future. Then the algorithm is examined using several experiments and benchmarks to test the performance through the corresponding computer program in the thereupon. Finally the Acknowledgments section expresses my gratitude to the help and the References section claims the referential resources used while conducting the project.

@@

	First, we need to observe what notation is needed by the construction. Apparently, we need to at least define these notations: \(\text{inf}\), \(\text{int}\), \(\text{num}\),\textit{  }\textit{ \(\sum\)}, \(=\), \({}^{\wedge}\), \(/\), \(-\), \(\text{abs}\), \(<\) and \(\text{for}\), where \(\text{abs}\) stands for the absolute value function. However, in this neat example, the concept of function is not needed to be defined. Additionally, as the constructed stuff only needs to demonstrate the equation itself instead of the principles behind the equality, there is no necessity for the definition of notation \(+\). Note that manifestly \(\text{int} \triangleleft | \text{num}\). The table below (ref{table-1}) demonstrates the number of parameters taken by the notations.

	table 1 here.

@@

	equations here.

@@

	equations here.

@@

		equations here.

@@

		Graphically, figure ref{figure-1} illustrates the initial map.

		figure 1 here.

@@

	algorithm 1 here.

@@

	A situation after some reinforcements may be the one demonstrated in figure ref{figure-2}.

	figure 2 here.

@@

	To summarize, we can illustrate the entire process of the workflow of the algorithm with the following diagram on the next page (figure ref{figure-3}).

	figure 3 here.

@@

	\section{Testing the algorithm's performance through benchmarks}

	The discussion below is based on an implementation of the algorithm by us, called DefQed and is open-sourced on \cite{Wang2022}. It will be further described in the sixth section.

	To illustrate the performance of the brand-new algorithm, we construct a relatively simple exemplification. The source code describing the demonstration can be found on the internet through \cite{Wang2022-2}. Alongside the benchmark code of the algorithm in the repository is the benchmark code of the popular theorem prover Microsoft Z3. The code for Wolfram Engine (or Wolfram Mathematica) is also presented.

	The timing cost by the three provers are shown as below in ref{table-2}.

	table 2 here.

	The data can be further illustrated in ref{figure-4}. The blue line stands for the time (milliseconds) of DefQed, while the orange line stands for the time (milliseconds) of Wolfram Engine.

	figure 4 here.

	From ref{figure-4} we can obviously tell the time cost of the reasoning increases by the round, as more reflections are inserted into the database. It should be noted that for time reasons, the reduction algorithm is not implemented in this version of DefQed, and the coding quality can also be further improved.

	The experiment is done on a Hewlett-Packard ZBook mobile workstation 14u G5 \cite{HP-2018}. During the experiment, all the programs/applications not necessary are closed except for the terminal emulator (ConEmu) and the program to benchmark. The operating system is Microsoft Windows 11 Pro for Workstations Insider Preview (Build 25211). The processor is Intel Core i5 8250U and the memory size is 8.0 GB installed. We used Wolfram Language version 12.0 and DefQed version 0.03. As DefQed utilizes MySQL as its database, it is necessary to point out we are using MySQL version 8.0 (Community). The DefQed source code is built with option `Release Mode' for Microsoft Windows x64 processor with .NET 6.0 platform.

	We also need to point out that though in each round DefQed seems to be slower than others, it does not mean that the algorithm has a worse efficiency and performance when comparing to others. The performance difference between platform and programming language utilized by the two, .NET and mainly C contributes to the difference of the two applications, not mentioning that while Wolfram Engine can only prove equations, DefQed is a generic approach. The quality of the optimization of the codes are also not on be same basis.

@@

	In the sections before we have presented an elegant data structure for representing mathematical concepts and a neat process which utilizes category theory concepts and knowledge to conduct proof, which also have self-optimization features.

@@

	\section{Acknowledgments}
	I express my sincerest gratitude to my tutor PSX, the Y Project of P.R.China, which enables senior high school students to conduct science research, my senior high academy NXXX of P.R.China and the corresponding university XXU of P.R.China. I also thank my parents for encouraging me to conduct this research project.

@@

	\bibliographystyle{unsrtnat}
	\bibliography{tot}

\end{document}
